Detection of Oscillations in Aldebaran with Ground-Based Observations

### 2017-12-22 presub/natastron_presub_abstract.tex
presubmission letter and abstract draft

% !TEX TS-program = pdflatexmk

\documentclass[modern]{aastex61}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{xspace}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathrsfs}

\include{macros}
% Journals

% Astronomical abbreviations (thanks to Dan Huber)
\newcommand{\numax}{\mbox{$\nu_{\rm max}$}\xspace}
\newcommand{\Dnu}{\mbox{$\Delta \nu$}\xspace}
\newcommand{\dnu}{\mbox{$\delta \nu$}\xspace}
\newcommand{\muHz}{\mbox{$\mu$Hz}\xspace}
\newcommand{\teff}{\mbox{$T_{\rm eff}$}\xspace}
\newcommand{\logg}{\mbox{$\log g$}\xspace}
\newcommand{\feh}{\mbox{$\rm{[Fe/H]}$}\xspace}
\newcommand{\msun}{\mbox{$\mathrm{M}_{\odot}$}\xspace}
\newcommand{\lsun}{\mbox{$\mathrm{L}_{\odot}$}\xspace}
\newcommand{\mearth}{\mbox{$\mathrm{M}_{\oplus}$}\xspace}
\newcommand{\rsun}{\mbox{$\mathrm{R}_{\odot}$}\xspace}
\newcommand{\kepler}{\emph{Kepler}\xspace}
\newcommand{\hipparcos}{\emph{Hipparcos}\xspace}
\newcommand{\gaia}{\emph{Gaia}\xspace}
% \newcommand{\ktwo}{\textit{K2}\xspace}
\newcommand{\ktwo}{\emph{K2}\xspace}
\newcommand{\ksc}{{\sc k2sc}\xspace}
\newcommand{\ksf}{{\sc k2sff}\xspace}
\newcommand{\kms}{\,km\,s$^{-1}$} % kilometres per second
\newcommand{\bibtex}{\textsc{Bib}\!\TeX} % bibtex. Not quite the correct typesetting, but close enough

\newcommand{\guy}[1]{{\bf \color{blue} #1}}

@@

\title{Detection of Oscillations in Aldebaran with Ground-Based Observations}

\author[0000-0003-1540-8562]{Will M. Farr}

\affiliation{Birmingham Institute for Gravitational Wave Astronomy,
  University of Birmingham, Birmingham, B15 2TT, United Kingdom}

\affiliation{School of Physics and Astronomy, University of
  Birmingham, Birmingham, B15 2TT, United Kingdom}

\author[0000-0002-4290-7351]{Guy R. Davies}

\affiliation{School of Physics and Astronomy, University of
  Birmingham, Birmingham, B15 2TT, United Kingdom}
\affiliation{Stellar Astrophysics Centre, Department of Physics 
and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000 Aarhus C, Denmark}

\author[0000-0003-2595-9114]{Benjamin J. S. Pope}
\affiliation{Sydney Institute for Astronomy, School of Physics, University of Sydney, Sydney NSW 2006, Australia} 
\affiliation{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, New York, NY 10003, USA}
\affiliation{NASA Sagan Fellow}

\author{Thomas North}
\affiliation{School of Physics and Astronomy, University of Birmingham, Birmingham, B15 2TT, United Kingdom}
\affiliation{Stellar Astrophysics Centre, Department of Physics 
and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000 Aarhus C, Denmark}

\author[0000-0002-6980-3392]{Timothy R. White}
\affiliation{Stellar Astrophysics Centre, Department of Physics 
and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000 Aarhus C, Denmark}

\author{James Barrett}
\affiliation{Birmingham Institute for Gravitational Wave Astronomy,
  University of Birmingham, Birmingham, B15 2TT, United Kingdom}

\affiliation{School of Physics and Astronomy, University of
  Birmingham, Birmingham, B15 2TT, United Kingdom}

\author{Andrea Miglio}

\affiliation{School of Physics and Astronomy, University of
  Birmingham, Birmingham, B15 2TT, United Kingdom}
\affiliation{Stellar Astrophysics Centre, Department of Physics 
and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000 Aarhus C, Denmark}

\author{Mikkel N. Lund}

\affiliation{Stellar Astrophysics Centre, Department of Physics 
and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000 Aarhus C, Denmark}
\affiliation{School of Physics and Astronomy, University of
  Birmingham, Birmingham, B15 2TT, United Kingdom}

\author{Victoria Antoci}
\affiliation{ Stellar Astrophysics Centre (SAC), Department of Physics and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000, Denmark}

\author{Mads Fredslund Andersen} 
\affiliation{ Stellar Astrophysics Centre (SAC), Department of Physics and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000, Denmark}
\author{Frank Grundahl}
\affiliation{ Stellar Astrophysics Centre (SAC), Department of Physics and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000, Denmark}

\email{w.farr@bham.ac.uk, G.R.Davies@bham.ac.uk, benjamin.pope@nyu.edu, txn016@student.bham.ac.uk, jimbarrett27@gmail.com}

\begin{abstract}
The nearby red giant Aldebaran is suspected to host a gas giant planetary companion from decades of ground-based spectroscopic measurements of its Doppler shift. Using Gaussian Process-based Continuous Auto-Regressive Moving Average (CARMA) models, we confirm this suspected planet and show that these historic data also contain evidence of acoustic oscillations in the star itself, and verify this result with further dedicated ground-based spectroscopy and space-based photometry with the \kepler Space Telescope. From the frequency of these oscillations we determine the mass of Aldebaran to be $1.16 \pm 0.07$~\msun, and note that this implies its planet will have been subject to insolation comparable to the Earth for some of the star's main sequence lifetime. Our approach to sparse, irregularly sampled time series astronomical observations has the potential to unlock similar measurements for hundreds of stars in archival data, and permits more flexible observing schedules in future. 
\end{abstract}

\section{Introduction}

Aldebaran, or $\alpha$~Tauri, is a well-known first-magnitude naked-eye red giant star, and has long been the subject of astronomical investigations. It was one of the first stars around which an extrasolar planet candidate was identified, by looking for Doppler shifts from the star's reflex motion around the common centre of mass with its companion \citep[the radial velocity or RV method;][]{struverv}. While the hot Jupiter 51~Peg~b \citep{51peg} was the first exoplanet to be recognized as such, before this, \citet{hatzes1993} had noted RV variations in Pollux \citep[$\beta$~Gem; subsequently confirmed as a planet:][]{betgemconf,betgemconf2}, Arcturus ($\alpha$~Boo, unconfirmed), and Aldebaran. After further investigation by \citet{Hatzes1998}, \citet{Hatzes2015} have now claimed a firm RV detection of a planetary-mass companion Aldebaran~b, with a period of $628.96 \pm 0.90$~d.

In this paper, we present a re-analysis of these original RV data in which we not only confirm this signal, but detect acoustic oscillations in Aldebaran for the first time. We validate this method and its result with new independent RV observations with the SONG~Telescope, and photometry from the \ktwo Mission. By measuring the frequency of maximum power, \numax, of these \emph{p}-mode oscillations we asteroseismically determine the mass of Aldebaran to be $1.16 \pm 0.07$~\msun. This measured stellar mass allows us to calculate that any moons of this giant planet, although they are now likely to be very hot, may have had equilibrium temperatures comparable to that of the Earth when Aldebaran was on the main sequence, raising the possibility that they may have once been habitable.

Our new approach to asteroseismic data analysis, based on Continuous Auto-Regressive Moving Average (CARMA) models, can extract exoplanet signals together with measures of \numax from sparse and irregularly-sampled time series. An all-sky survey to find planetary companions and to precisely measure the masses of all nearby red giant stars is feasible with this new approach, and the required data either already exist in large radial velocity exoplanet surveys, or are easy to obtain with ground-based telescopes.

\section{Asteroseismology of Red Giants}
Asteroseismology is a powerful tool for the charaterisation of red giant stars \citep[see][for a detailed review]{2013ARA&A..51..353C}.  Red giants are solar-type pulsators, they exhibit acoustic modes of oscillation that are both excited and damped by stellar convection.  In the frequency-power spectrum of either radial velocity of photometric observations, these modes of oscillation form an often intricate pattern of sharp peaks where the properties of the pattern can be mapped back to intrinsic stellar properties \citep[e.g.,][]{2016AN....337..774D}.  The easiest properties of the pattern to determine are the the frequency of maximum amplitude \numax and the so-called large separation \Dnu \citep{Kjeldsen95}.  The two observables are often referred to as the global average seismic parameters.

Stellar properties can be determined with the average seismic measures and effective temperature \teff.  A set of scaling relations exist that illustrate the constraint the global asteroseismic measures provide.  These scaling relations are understood to be approximate but never the less are commonly used for radius $R$, mass $M$, and surface gravity $g$:
\begin{eqnarray}
\left( \frac{R}{\mathrm{R_{\odot}}} \right) \simeq & \left( \frac{\nu_{\rm max}}{\nu_{\rm max, \odot}} \right) \, 
\left( \frac{\Delta \nu}{\Delta \nu_{\odot}} \right)^{-2} \, \left( \frac{T_{\rm eff}}{T_{\rm eff, \odot}} \right)^{0.5}, \\
\left( \frac{M}{\mathrm{M_{\odot}}} \right) \simeq & \left( \frac{\nu_{\rm max}}{\nu_{\rm max, \odot}} \right)^{3} \, 
\left( \frac{\Delta \nu}{\Delta \nu_{\odot}} \right)^{-4} \, \left( \frac{T_{\rm eff}}{T_{\rm eff, \odot}} \right)^{1.5}, \\
\left( \frac{g}{\mathrm{g_{\odot}}} \right) \simeq & \left( \frac{\nu_{\rm max}}{\nu_{\rm max, \odot}} \right) \, 
 \left( \frac{T_{\rm eff}}{T_{\rm eff, \odot}} \right)^{0.5}. \\
\end{eqnarray}
While the accuracy of these scaling relations is still a matter of ongoing work \citep[e.g.,][]{2017ApJ...844..102H}, stellar parameters can be estimated by comparing observables to parameter from models of stellar evolution.

As can be seen from the scaling relations, asteroseismology can provide constraint on Mass, Radius, and surface gravity if \Dnu, \numax, and \teff can be measured.  In practise, for the higher luminosity red giants it is more straight forward to measure \numax than \Dnu.  Typical values for \numax might range from $0.1$ to $20 \rm\,\mu Hz$ for the luminous giants and $20$ to $50 \rm\,\mu Hz$ for red giants with luminosities similar to the red clump .  For the same sets of stars, \Dnu would take ranges from $0.02$ to $3 \rm\,\mu Hz$ and $3$ to $7 \rm\,\mu Hz$ depending on stellar mass and \teff (or metallicity which is considered more fundamental) \citep[e.g.,][]{2011A&A...525L...9M, 2013A&A...559A.137M}.  Hence, \numax requires less frequency resolution than \Dnu to establish a good measurement.

The ability to measure \numax and \Dnu, therefore, depends on the length and sampling rate of a data set.  For regularly spaced data, typically analysed using an estimate of the frequency-power spectrum, the Nyquist limit (half the sampling rate) and frequency resolution ($\sim 1/$~duration) are trivial to estimate. %For the historic data sets proposed here the typical length is \guy{blah} and typical sampling is \guy{blah}.  This in principle gives us access to measures of \numax for \guy{these types of stars} and \Dnu $\,$ \guy{for those types of stars}.

\section{Time-Domain Models}

It is easy to observe Aldebaran and similarly-bright stars with ground-based spectroscopic instruments, typically only requiring short exposures that can be obtained even under adverse observing conditions. There is indeed a considerable archive of such observations already, as a legacy of radial velocity (RV) surveys conducted to find exoplanets. In most cases, however, these have not so far been useful for asteroseismology, because these RV data are sparsely and irregularly-sampled. Because we have to pause observations during the day, during poor weather conditions, or simply when targets of higher priority are being observed, we get time series of only a few points and which may have significant and uneven gaps. This introduces a window-function effect: the power spectrum as constructed for example by a Fourier transform, or a Lomb-Scargle Periodogram \citep{lomb,scargle}, is convolved with the Fourier transform of the window function, introducing strong sidelobes adjacent to real frequency peaks and causing crosstalk between adjacent frequency channels. This imposes significant limitations both on the signal-to-noise and frequency resolution of power spectra derived from linear methods such as the Lomb-Scargle periodogram, and in practice makes asteroseismology with these conventional approaches difficult or impossible from the ground for stars with oscillation periods ranging from $\sim 12$~h to $\sim$~a few days.

If we apply nonlinear statistical inference methods this situation can be improved. \citet{brewer2009} show that a system of driven, damped harmonic oscillators such as we encounter in asteroseismology can be statistically modelled as a Gaussian Process, with a covariance kernel consisting of a sum of damped sinusoids. The hyperparameters of this process can be used to retrieve very accurate power spectral density distributions. Unfortunately, this is impractical for long time series, as the computational cost of evaluating this likelihood function scales as $\mathcal{O}(n^3)$.

Reference \citet{Kelly2014}, compare to \citet{Foreman-Mackey2017}.

\section{Aldebaran}
Aldebaran is a red giant star with spectral type K5, one of the nearest such stars at a distance of only $19.96 \pm 0.38$~pc \citep{hipparcos}. Its position near the Ecliptic permits the determination of its angular diameter by lunar occultations and by interferometry \citep[$20.58 \pm 0.03$ mas;][]{richichi2005,1979ApJ...228L.111B,brown1979,panek1980}.

\subsection{SONG Observations}    % this goes into section 4.1

In order to detect the oscillations of Aldebaran via the radial velocity 
method we used the Hertzsprung SONG telescope \cite{2017ApJ...836..142G}. All 
observations were carried out in the highest resolution mode (R$\approx\,110.000$)
with an integration time of 30s. The observations were carried out between 27 September 
2016 and  12 January 2017, during this time we attempted to obtain at least one 
visit per available night. A total of 254 spectra was obtained. The spectra were
extracted using the SONG pipeline (see \cite{2017ApJ...836..142G}). SONG employs an iodine cell for
precise wavelength calibration and the velocities were determined using the $\it iSONG$
software following the same procedures as described in \cite{2017ApJ...836..142G}.   The typical
velocity precision per visit is 2.5$m\,s^{-1}$ allowing us to easily detect the oscillations
which displays a peak-to-peak amplitude of $\approx \, 170 \, \rm m \, s^{-1}$. $\bf Figure XX$ show
the radial velocities obtained -- oscillations are easily visible as well as a long-term
trend. This is probably due to the a manifestation of the planetary companion reported by
\cite{Hatzes2015}.

\subsection{K2 Observations}

%In order to verify the results of the novel analysis presented above, we sought to obtain an independent detection of the oscillations of Aldebaran and compare the frequencies determined with the two methods. 
Aldebaran was observed with \ktwo \citep{howell14}, a two-wheeled revival of the \kepler Mission \citep{2010sci...327..977b}, under Guest Observer Program 130471 in Campaign~13, from 2017-03-08 to 2017-05-27. Aldebaran and similar high-luminosity K giants are well known to show photometric variability due to solar-like oscillations \citep{bedding2000}, and hence the continuous, high-precision K2 light curves allow an independent confirmation of our CARMA results.

As Aldebaran is extremely bright, it saturates the \kepler detector and it is therefore not possible to use standard photometry pipelines to extract a K2 lightcurve. We instead use halo photometry \citep{White2017}, whereby unsaturated pixels from the outer part of the large and complicated halo of scattered light around bright stars are used to reconstruct a light curve as a weighted linear combination of individual pixel-level time series, as described in Appendix~\ref{halo}.

Figure \ref{k2_lightcurve} shows the light curve and power spectrum of the K2 observations of Aldebaran. We detect clear variability on a average timescale of $\sim$\,5\,days, consistent with the expected timescale for solar-like oscillations for a high-luminosity red giant. To measure global asteroseismic parameters we model the background variability in the Fourier domain using the methodology described in \citet{huber09}, yielding a frequency of maximum power of $\numax = 2.2 \pm 0.25\,\muHz$ and amplitude per radial mode of $1850 \pm 500$\,ppm. Due to the limited frequency resolution of the 70-day K2 timeseries we were unable to measure the large frequency separation \Dnu, which is expected to be on the order of $\approx$\,0.5\muHz.

Using the \textsc{k2ps} planet-search code \citep{k2ps,Pope2016} to examine this light curve, we search for transits across a wide range of periods, and find no evidence either of short-period planetary transits or an eclipsing stellar companion.

\begin{figure}
\centering
% \includegraphics[width=0.75\textwidth]{Aldebaran_lc.png}
%\includegraphics[width=0.75\textwidth]{aldebaran_lc.eps}
\includegraphics[width=\textwidth]{k2obs.png}
\caption{\ktwo lightcurve (left) and power spectrum (right) of Aldebaran. The red dashed line shows the background model, and the orange line is a heavily smoothed version of the power spectrum used to measure the frequency of maximum power. }
\label{k2_lightcurve}
\end{figure}

\subsection{Planetary and Stellar Parameters}

By jointly fitting a Keplerian for the planet with a CARMA model for the stellar oscillations, we obtain a precise estimate of both \numax and the planetary orbital parameters. The planetary orbital parameters we obtain are similar to \citet{Hatzes2015}: period $P = 629.0^{+ 1.7}_{-2.1}$~d, eccentricity $e = 0.168^{+ 0.069}_{- 0.071}$; and radial velocity semi-amplitude $K = 127^{+ 15}_{- 14}$~m/s.

The stellar oscillations are best fit with a single CARMA component with $\numax = 2.238^{+ 0.086}_{- 0.080}$ \muHz, and including additional CARMA components does not resolve any further structure in the power spectrum. We have determined the stellar properties of Aldebaran with the additional constraint of \numax determined from this analysis.  The details of the stellar modelling are presented in detail in \ref{sec:mod}.  We have considered the impact of our additional \numax constraint and run our analysis for different sets of spectroscopic estimates.  For final values we adopt the \citet{2012Sheffield} spectroscopic solution as being `middle of the road' estimates.  For the analysis without constraint on \numax we find estimates of the stellar properties as $M = 1.27^{+0.24}_{-0.2} \, \mathrm{M_{\odot}}$ and age $4.86^{+3.56}_{-2.04} \, \rm Gyr$.  With the inclusion of \numax we find $M = 1.16^{+0.07}_{-0.07} \, \mathrm{M_{\odot}}$ and age $6.38^{+1.42}_{-1.12} \, \rm Gyr$.  It is clear that the addition of \numax provides substantially more precise estimates of mass and age. With this stellar mass, the radial velocity signal translates to an inclination-dependent planetary mass of $M\,\sin\,{i} = 5.81 \pm 0.72~M_J$ \citep[Equation~1]{Torres2008}, a massive giant planet; for sufficiently high inclinations ($\gtrsim 45^\circ$), the mass of Aldebaran~b may exceed $13~M_J$, making it a possible low-mass brown dwarf candidate.  % re sheffield - I have substituted the citation from the .bib but you said 'b'; afaik she only published one paper in 2012? - Ben

With the determination of the stellar mass it is possible to infer the parameters Aldebaran had while it was on the main sequence. We conduct a Monte Carlo simulation, drawing masses randomly from a Gaussian distribution $M \sim \mathcal{N}(1.16,0.07)$ and metallicities $[\text{Fe}/\text{H}] \sim \mathcal{N}(-0.15,0.2)$ \citep{decin2003} to predict the luminosity of the main-sequence progenitor, and the semi-major axis of the planet's orbit.  %Using the scaling relations of \citep{eker2015}, we find that Aldebaran would have had a luminosity $L_* = 2.0 \pm 0.37~\lsun$ on the main sequence.  
Using the Mesa Isochrones and Stellar Tracks \citep[MIST:][]{mist0,mist1} models via the \textsc{isochrones} package \citep{isochrones}, we compute the evolution of the luminosity $L$ of the star along the main sequence; $L$ evolves from $2.0 \pm 0.7$~\lsun at 0.5~Gyr to $3.5 \pm 2.3$~\lsun at 4.5~Gyr. 
We note that this implies that the planet at a semi-major axis of $1.50\, \pm\, 0.03$~AU would have been subject to an insolation comparable to that of the Earth, evolving from 0.5~Gyr to 4.5~Gyr from $0.86\, \pm\, 0.27$ to $1.55 \,\pm\,1.0$ times that of the Earth today. Subject as well as this to the great uncertainties of the planet's orbital evolution and albedo, Aldebaran~b and any of its moons may well have hosted temperate environments for some of their history, now long-since destroyed by their star's evolution away from the main sequence.

\section{Conclusions}

With CARMA models we have confirmed the previously-suspected planet around Aldebaran using archival data, and additionally detected acoustic oscillations that permit a $\sim 6\%$ mass determination. We have confirmed both of these results with new ground-based observations, and space-based photometry with \ktwo.

From this pilot study with Aldebaran, have shown that with limited quantities of data of limited quality, we can nevertheless do precision asteroseismology from the ground and obtain very precise estimates of stellar parameters. Furthermore, improvements in the algorithm may also permit the detection of not only~\numax but also~\Dnu from similar measurements, permitting more sophisticated asteroseismic analysis. Sufficient radial velocity data either already exist, or are trivial to obtain, in order to do this for essentially all bright giants; while for solar-like stars, these new methods allow significantly relaxed observing requirements. One rich archive is the series of observations with the Hamilton \'{E}chelle Spectrograph at Lick Observatory, studying 373 G~and K~giants \citep[e.g. ][]{frink2001,hekker2006,ortiz2016}. The method may also be useful in detecting rotational or other periodic variability in the LCES HIRES/Keck Precision Radial Velocity Exoplanet Survey \citep{butler17} of 1,624 FGKM dwarfs, although the time sampling is by design too coarse to detect $p$-mode oscillations in these stars. Furthermore, CARMA models and related methods will enhance deep, all-sky, sparse photometric surveys: an immediate future test of this will be from \hipparcos; while only~58 epochs of photometry are available for Aldebaran, at a sampling we find to be insufficient for our purposes, stars at higher latitudes may often have 150--200 epochs and more even sampling \citep{hipparcos_phot}, and the~14 bright~K giants in \hipparcos noted by \citet{bedding2000} are an ideal first test case. This can naturally be extended to \emph{Gaia} in space \citep{gaia}, or \emph{LSST} from the ground \citep{dmt,lsst,lsstbook}, from which many thousands of new asteroseismic determinations will be possible.

\acknowledgments

We thank the academy....

We would like to thank Tim Bedding, David Hogg and Dan Foreman-Mackey for their helpful discussions.

BP is grateful for funding from the Breakthrough Prize Foundation. This work was performed in part under contract with the Jet Propulsion Laboratory (JPL) funded by NASA through the Sagan Fellowship Program executed by the NASA Exoplanet Science Institute.

The authors would like to acknowledge the Gadigal people of the Eora nation on whose ancestral lands the University of Sydney is built.

Funding for the Stellar Astrophysics Centre is provided by The Danish National Research Foundation. TRW acknowledges the support of the Villum Foundation (research grant 10118).
This research has made use of the SIMBAD database, operated at CDS, Strasbourg, France, and NASA's Astrophysics Data System. The \kepler data presented in this paper were obtained from the Mikulski Archive for Space Telescopes (MAST). STScI is operated by the Association of Universities for Research in Astronomy, Inc., under NASA contract NAS5-26555. Support for MAST for non-HST data is provided by the NASA Office of Space Science via grant NNX13AC07G and by other grants and contracts.

The Villum Foundation, The Danish Council for Independent Research~|~Natural Sciences and 
 the Carlsberg Foundation are thanked for their support on building the SONG prototype on 
 Tenerife. 
 We thank Jens Jessen-Hansen, Rasmus Handberg and René Tronsgaard Rasmussen for their contribution
 to the development of the SONG data reduction pipeline and archive system.

Based on observations made with the Hertzsprung SONG telescope operated on the Spanish Observatorio
  del Teide on the island of Tenerife by the Aarhus and Copenhagen Universities and by the Instituto de 
  Astrofisica de Canarias.

This work made use of the \textsc{IPython} package \citep{PER-GRA:2007}; SciPy \citep{scipy};  \textsc{matplotlib}, a \textsc{Python} library for publication quality graphics \citep{Hunter:2007}; \textsc{Astropy}, a community-developed core \textsc{Python} package for astronomy \citep{2013A&A...558A..33A}; and \textsc{NumPy} \citep{van2011numpy}. Some of the acknowledgements were compiled using the Astronomy Acknowledgement Generator.

\appendix

\section{Halo Photometry}
\label{halo}

The \kepler Space Telescope \citep{2010sci...327..977b}  
suffered a critical reaction wheel failure in May 2013, which made it impossible to maintain a stable pointing and therefore continue its nominal mission. It was revived as \ktwo \citep{howell14}, balanced by orienting perpendicular to the Sun. This requires that \ktwo observes fields in the Ecliptic in $\sim~80$~d Campaigns.

The \kepler detector saturates for stars brighter than the $\sim~11^\text{th}$ magnitude. Nevertheless, the excess flux deposited in a saturated pixel spills conservatively up and down the pixel column, such that it is possible to sum this `bleed column' for bright stars and still obtain precise photometry, such as was done for the brightest star in the nominal \kepler mission, $\theta$~Cyg \citep[$V = 4.48$;][]{guzik2011,thetacygwhite,guzik2016}. There are two main reasons why this is not possible for all bright stars in general. First, because the downlink bandwidth from \kepler is very limited, it is often not desirable to download the large pixel stamps as are required for such bright stars; these must be even larger in \ktwo to account for the reduced image stability. Second, if the bleed column for a sufficiently bright star reaches the edge of the chip, flux spills over and is not conserved, imposing a hard brightness limit that depends on the distance to the detector edge. 
Collateral `smear' data, which are collected to help calibrate the photometric bias from stars sharing the same column as a target, can be used to reconstruct light curves for un-downloaded bright stars and thereby avoid bandwidth constraints \citep{k2smear}, but these data are still rendered unusable if the bleed column falls off the edge of the chip and contaminates the smear rows.

>>>>

\documentclass{natureprintstyle}
%\documentclass{nature}
\bibliographystyle{naturemag}
\usepackage{epsfig,caption}
\usepackage{color}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{amssymb}
\usepackage{rotating}
\usepackage{latexsym}
\usepackage{hyperref}
\usepackage{hyperref}
% \usepackage{natbib}
\usepackage{xspace}
\usepackage{amsmath}
\usepackage{mathrsfs}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%journal commands
\newcommand{\apj}{Astrophys. J.}
\newcommand{\spie}{Proc. SPIE}
\newcommand{\pasp}{Publ. Astron. Soc. Pac.}
\newcommand{\apjs}{Astrophys. J. Supp.}
\newcommand{\araa}{Annu. Rev. Astron. Astrophys.}
\newcommand{\mnras}{Mon. Not. R. Astron. Soc.}
\newcommand{\apjl}{Astrophys. J. Let.}
\newcommand{\aap}{Astron. Astrophys.}
\newcommand{\aj}{Astron. J.}
\newcommand{\nat}{Nature}
\newcommand{\na}{New Astron. Rev.}
\newcommand{\aaps}{A\&AS}
\newcommand{\procspie}{Proc. SPIE}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Astronomical abbreviations (thanks to Dan Huber)
\newcommand{\numax}{\mbox{$\nu_{\rm max}$}\xspace}
\newcommand{\Dnu}{\mbox{$\Delta \nu$}\xspace}
\newcommand{\dnu}{\mbox{$\delta \nu$}\xspace}
\newcommand{\muHz}{\mbox{$\mu$Hz}\xspace}
\newcommand{\teff}{\mbox{$T_{\rm eff}$}\xspace}
\newcommand{\logg}{\mbox{$\log g$}\xspace}
\newcommand{\feh}{\mbox{$\rm{[Fe/H]}$}\xspace}
\newcommand{\msun}{\mbox{$\mathrm{M}_{\odot}$}\xspace}
\newcommand{\lsun}{\mbox{$\mathrm{L}_{\odot}$}\xspace}
\newcommand{\mearth}{\mbox{$\mathrm{M}_{\oplus}$}\xspace}
\newcommand{\rsun}{\mbox{$\mathrm{R}_{\odot}$}\xspace}
\newcommand{\kepler}{\emph{Kepler}\xspace}
\newcommand{\hipparcos}{\emph{Hipparcos}\xspace}
\newcommand{\gaia}{\emph{Gaia}\xspace}
% \newcommand{\ktwo}{\textit{K2}\xspace}
\newcommand{\ktwo}{\emph{K2}\xspace}
\newcommand{\ksc}{{\sc k2sc}\xspace}
\newcommand{\ksf}{{\sc k2sff}\xspace}
\newcommand{\kms}{\,km\,s$^{-1}$} % kilometres per second
\newcommand{\bibtex}{\textsc{Bib}\!\TeX} % bibtex. Not quite the correct typesetting, but close enough

\newcommand{\guy}[1]{{\bf \color{blue} #1}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% title brainstorm:
% According to nature astronomy style guide: "a title of, at most, 90 characters (including spaces)"
%formal document says 100 characters
\title{Detection of Oscillations in Aldebaran with Ground-Based Observations}

\author{Will M. Farr$^{1,2}$\thanks{w.farr@bham.ac.uk}, Guy R. Davies$^{2,3}$, Benjamin J. S. Pope$^{4,5,6}$, et al.}

@@

\maketitle

\let\thefootnote\relax\footnote{
\begin{affiliations}
\item {Birmingham Institute for Gravitational Wave Astronomy,
  University of Birmingham, Birmingham, B15 2TT, United Kingdom}
\item {School of Physics and Astronomy, University of
  Birmingham, Birmingham, B15 2TT, United Kingdom}
\item {Stellar Astrophysics Centre, Department of Physics 
and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000 Aarhus C, Denmark}
\item {Sydney Institute for Astronomy, School of Physics, University of Sydney, Sydney NSW 2006, Australia} 
\item {Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, New York, NY 10003, USA}
\item {NASA Sagan Fellow}

\end{affiliations}
}

\begin{abstract}

The nearby red giant Aldebaran is suspected to host a gas giant planetary companion from decades of ground-based spectroscopic measurements of its Doppler shift. Using Gaussian Process-based Continuous Auto-Regressive Moving Average (CARMA) models, we confirm this suspected planet and show that these historic data also contain evidence of acoustic oscillations in the star itself, and verify this result with further dedicated ground-based spectroscopy and space-based photometry with the \kepler Space Telescope. From the frequency of these oscillations we determine the mass of Aldebaran to be $1.16 \pm 0.07$~\msun, and note that this implies its planet will have been subject to insolation comparable to the Earth for some of the star's main sequence lifetime. Our approach to sparse, irregularly sampled time series astronomical observations has the potential to unlock similar measurements for hundreds of stars in archival data, and permits more flexible observing schedules in future.

Aldebaran, or $\alpha$~Tauri, is a well-known first-magnitude naked-eye red giant star, and has long been the subject of astronomical investigations. It was one of the first stars around which an extrasolar planet candidate was identified, by looking for Doppler shifts from the star's reflex motion around the common centre of mass with its companion\cite{struverv}. While the hot Jupiter 51~Peg~b was the first exoplanet to be recognized as such in 1995\cite{51peg}, as early as 1993\cite{hatzes1993} RV variations in Pollux \cite{betgemconf,betgemconf2}, Arcturus (as yet unconfirmed), and Aldebaran suggested these giant stars may themselves host massive planets. After further investigation\cite{Hatzes1998}, Hatzes et al.\cite{Hatzes2015} have now claimed a firm RV detection of a planetary-mass companion Aldebaran~b, with a period of $628.96 \pm 0.90$~d.

In this paper, we present a re-analysis of these original RV data in which we not only confirm this signal, but detect acoustic oscillations in Aldebaran for the first time. We validate this method and its result with new independent RV observations with the SONG~Telescope, and photometry from the \ktwo Mission (displayed in Figure~\ref{fig:k2obs}. By measuring the frequency of maximum power of these \emph{p}-mode oscillations we asteroseismically determine the mass of Aldebaran to be $1.16 \pm 0.07$~\msun. This measured stellar mass allows us to calculate that any moons of this giant planet, although they are now likely to be very hot, may have had equilibrium temperatures comparable to that of the Earth when Aldebaran was on the main sequence, raising the possibility that they may have once been habitable.

Our new approach to asteroseismic data analysis, based on Continuous Auto-Regressive Moving Average (CARMA) models, can extract exoplanet signals together with measures of the frequency of maximum power from sparse and irregularly-sampled time series. An all-sky survey to find planetary companions and to precisely measure the masses of all nearby red giant stars is feasible with this new approach, and the required data either already exist in large radial velocity exoplanet surveys, or are easy to obtain with ground-based telescopes.

\end{abstract}

\begin{figure}[!t]
\begin{center}
\includegraphics[width=\linewidth]{k2obs.png}
\caption{\ktwo lightcurve (left) and power spectrum (right) of Aldebaran. The red dashed line shows the background model, and the orange line is a heavily smoothed version of the power spectrum used to measure the frequency of maximum power. }
\label{fig:k2obs}
\end{center}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\bibliographystyle{mnras}
\bibliography{aldebaran.bib}

%\begin{thebibliography}{10}
%\expandafter\ifx\csname url\endcsname\relax
%  \def\url#1{\texttt{#1}}\fi
%\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
%\providecommand{\bibinfo}[2]{#2}
%\providecommand{\eprint}[2][]{\url{#2}}
%
%\bibitem{TNaab2009}
%\bibinfo{author}{{Naab}, T.}, \bibinfo{author}{{Johansson}, P.~H.} \&
%  \bibinfo{author}{{Ostriker}, J.~P.}
%\newblock \bibinfo{title}{{Minor Mergers and the Size Evolution of Elliptical
%  Galaxies}}.
%\newblock \emph{\bibinfo{journal}{\apjl}} \textbf{\bibinfo{volume}{699}},
%  \bibinfo{pages}{L178--L182} (\bibinfo{year}{2009}).
%
%\end{thebibliography}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
